Data Cleaning

### 2025-12-08 Data Cleaning.Rmd
Updated alkalinity/hardness cleaning to fix error that did not remove outliers.
Edited lag addition. Added a dose change column for each of coagulant doses to run predictive model on dose change.

merged_clean2 <- merged[merged$alkalinity < 876, ] # remove outliers

@@

```{r}
data <- read.csv("RWAR_dosing_cleaned.csv")
lagged_data <- data |>
  mutate(Temp_lag1 = lag(mean_Temp, 1), 
         Temp_lag2 = lag(mean_Temp, 2),
         pH_lag1 = lag(mean_pH, 1),
         pH_lag2 = lag(mean_pH, 2),
         Cond_lag1 = lag(mean_Conductivity, 1),
         Cond_lag2 = lag(mean_Conductivity, 2),
         Turb_lag1 = lag(mean_Turbidity, 1),
         Turb_lag2 = lag(mean_Turbidity, 2))
# Temp
lagged_data$Temp_lag1[1] <- lagged_data$Temp_lag1[2]
lagged_data$Temp_lag2[1:2] <- lagged_data$Temp_lag2[3]
# pH
lagged_data$pH_lag1[1] <- lagged_data$pH_lag1[2]
lagged_data$pH_lag2[1:2] <- lagged_data$pH_lag2[3]
# Conductivity
lagged_data$Cond_lag1[1] <- lagged_data$Cond_lag1[2]
lagged_data$Cond_lag2[1:2] <- lagged_data$Cond_lag2[3]
# Turbidity
lagged_data$Turb_lag1[1] <- lagged_data$Turb_lag1[2]
lagged_data$Turb_lag2[1:2] <- lagged_data$Turb_lag2[3]

lm_fit <- lm(Ferric.chloride..mg.L ~ ., data=lagged_data |> select(-c(Timestamp)))
summary(lm_fit)

@@

>>>>

merged_clean2 <- merged_clean1[merged_clean1$alkalinity < 120, ] # remove outliers

@@

```{r}
data <- read.csv("RWAR_dosing_cleaned.csv")
rwar_dosing_lag <- data |>
  mutate(Temp_lag1 = lag(mean_Temp, 1), 
         Temp_lag2 = lag(mean_Temp, 2),
         pH_lag1 = lag(mean_pH, 1),
         pH_lag2 = lag(mean_pH, 2),
         Cond_lag1 = lag(mean_Conductivity, 1),
         Cond_lag2 = lag(mean_Conductivity, 2),
         Turb_lag1 = lag(mean_Turbidity, 1),
         Turb_lag2 = lag(mean_Turbidity, 2),
         FeCl_change = Ferric.chloride..mg.L - lag(Ferric.chloride..mg.L),
         Polymer_change = Cationic.Polimer..mg.L - lag(Cationic.Polimer..mg.L))

# Get rid of first two observations (have NA values now)
rwar_dosing_lag1 <- rwar_dosing_lag[3:nrow(lagged_data),]

class_df <- rwar_dosing_lag1

threshold_smallchange <- 0.05

# normal data
class_df$FeCl_change <- ifelse(
    test = abs(class_df$FeCl_change) < threshold_smallchange,
    yes = "same",
    no = ifelse(
        test = class_df$FeCl_change < 0,
        yes = "dec",
        no = "inc"
    )
)

# polymer data
class_df$Polymer_change <- ifelse(
    test = abs(class_df$Polymer_change) < threshold_smallchange,
    yes = "same",
    no = ifelse(
        test = class_df$Polymer_change < 0,
        yes = "dec",
        no = "inc"
    )
)

write_csv(rwar_dosing_lag1, )
# # Temp
# lagged_data$Temp_lag1[1] <- lagged_data$Temp_lag1[2]
# lagged_data$Temp_lag2[1:2] <- lagged_data$Temp_lag2[3]
# # pH
# lagged_data$pH_lag1[1] <- lagged_data$pH_lag1[2]
# lagged_data$pH_lag2[1:2] <- lagged_data$pH_lag2[3]
# # Conductivity
# lagged_data$Cond_lag1[1] <- lagged_data$Cond_lag1[2]
# lagged_data$Cond_lag2[1:2] <- lagged_data$Cond_lag2[3]
# # Turbidity
# lagged_data$Turb_lag1[1] <- lagged_data$Turb_lag1[2]
# lagged_data$Turb_lag2[1:2] <- lagged_data$Turb_lag2[3]

lm_fit <- lm(Ferric.chloride..mg.L ~ ., data=lagged_data1 |> select(-c(Timestamp)))
summary(lm_fit)

@@

### Add inc/dec/same for dose

class_df <- lagged_data1

# normal data
class_df$FeCl_change <- ifelse(
    test = abs(class_df$FeCl_change) < 0.01,
    yes = "same",
    no = ifelse(
        test = class_df$FeCl_change < 0,
        yes = "dec",
        no = "inc"
    )
)

# polymer data
class_df$Polymer_change <- ifelse(
    test = abs(class_df$Polymer_change) < 0.01,
    yes = "same",
    no = ifelse(
        test = class_df$Polymer_change < 0,
        yes = "dec",
        no = "inc"
    )
)

class_df
```
